---
title: "R Lab Final"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
```



```{r, package-loading, echo=FALSE, message=T, warning=FALSE}
library("dplyr")
library("RSQLite")
library("nycflights13") 
library(dplyr)
library(magrittr)
library(ggplot2)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
final_db <- nycflights13_sqlite()


flights_sqlite <- tbl(final_db, "flights")
weather_sqlite <- tbl(final_db, "weather")
airports_sqlite <- tbl(final_db, "airports")
planes_sqlite <- tbl(final_db, "planes")
airlines_sqlite<-tbl(final_db, "airlines")

flights<- data.frame(flights_sqlite, header=TRUE)
weather<- data.frame(weather_sqlite, header=TRUE)
airport <- data.frame(airports_sqlite, header=TRUE)
planes<- data.frame(planes_sqlite, header=TRUE)
airlines<- data.frame(airlines_sqlite, header=TRUE)

flights = tbl(final_db, "flights") %>% 
  collect() %>%
  mutate(canceled = is.na(arr_time))

flights$cancel<-ifelse(flights$canceled==TRUE, 1, 0)

fp_join <- left_join((flights), (planes), by = "tailnum")
fpa<-left_join(fp_join, airport, by= c("dest"="faa"))

flights<-flights %>%
  mutate(
    date = as.Date(paste(month, year,hour,sep="."),format = "%m.%Y.%d.%H")
  )
weather_join<-left_join(flights, weather, by=c("year","month","day","hour"))
```

##Indicating relationship between weather and departure delays
```{r}
fws<- ggplot(weather_join,
             aes(x=visib, y=dep_delay))

fws + geom_point(aes(alpha = 0.01, colour = humid, size = canceled))
```
Here we find that visbility increases the occurence of departure delays. There also seems to be a stronger correlation with low humidity and delays over high humidity and delays


```{r, echo=F}
fpa$actual_delay<-ifelse((fpa$dep_delay>0), fpa$dep_delay, NA)

delay_hour <- fpa %>%
  dplyr::group_by(hour) %>%
  dplyr::summarise(
    max_delay = max(actual_delay, na.rm=TRUE),
    avg_delay = mean(actual_delay, na.rm=TRUE),
    min_delay = min(actual_delay,na.rm=TRUE),
    count = n()
  )
```

##Indicating relationship between hour and departure delays, Max delays and Avg delays

```{r, echo=FALSE}
plot(delay_hour$max_delay, type='l', main="Delay Statistics by the Hour", xlab="hour",ylab="delay")
lines(delay_hour$avg_delay, col = "blue", lty="dashed")
```
This plot shows that early morning hours typically experience a delay, ranging from 200 to 
400 minutes if we reference the average (blue dashed line). After about the
5th or 6th hour of the day, average delays remain very low, increasing 
ever so slightly as the day goes on. The extreme flight delays typically occur between the 
6th and 12th hours of the day. 
```{r, echo=F}
delay_day <- fpa %>%
  dplyr::group_by(day) %>%
  dplyr::summarise(
    max_delay = max(actual_delay, na.rm=TRUE),
    avg_delay = mean(actual_delay, na.rm=TRUE),
    min_delay = min(actual_delay,na.rm=TRUE),
    count = n()
  )
```
##Indicating relationship between days and departure delays, Max Delays
```{r, echo=FALSE}
delay_day <- fpa %>%
  dplyr::group_by(day) %>%
  dplyr::summarise(
    max_delay = max(actual_delay, na.rm=TRUE),
    avg_delay = mean(actual_delay, na.rm=TRUE),
    min_delay = min(actual_delay,na.rm=TRUE),
    count = n()
  )
```
This plot only shows max delays for days throughout the given month. There is typically a spike in delays with in the 8 to 10 day mark, and then max delays typically die down 
as the month goes by.
```{r, echo=FALSE}
plot(delay_day$max_delay, type='l', main="Delay Max by the Day", xlab="day",ylab="delay")
```

#Indicating relationship between days and delays, Avg Delays
```{r, echo=FALSE}
plot(delay_day$avg_delay, col = "blue", type = 'l', main="Delay Avg by the Day", xlab="day",ylab="delay")
```
This represents average delays on a given day and shows very similar behavior to max delays
```{r, echo=FALSE}
delay_year <- fpa %>%
  dplyr::group_by(year.x, month, day) %>%
  dplyr::summarise(
    max_delay = max(actual_delay, na.rm=TRUE),
    avg_delay = mean(actual_delay, na.rm=TRUE),
    min_delay = min(actual_delay,na.rm=TRUE),
    count = n()
  )
```

##Indicating relationship between Time of Year and delay, Max Delays
```{r, echo=FALSE}
plot(delay_year$max_delay, type='l', main="Delay Max by the Time of Year", xlab="Time of Year",ylab="delay")
```
From this we can interpret that there is no significant trend in max delays 
during a given year. 

#Indicating relationship between Time of Year and delay, Avg Delays
```{r, echo=FALSE}
plot(delay_year$avg_delay, type="l",col="red", main="Average Delay by the Time of Year", xlab="Time of Year", ylab="delay")
```
The same can be concluded for average delays

```{r, echo=FALSE}
h_cancel<-glm(flights$canceled~flights$hour, family=binomial(link="probit"))
d_cancel<-glm(flights$canceled~flights$day, family=binomial(link="probit"))
m_cancel<-glm(flights$canceled~flights$month, family=binomial(link="probit"))
```


#Indicating Relationship between time and cancelation
```{r, echo=FALSE}
summary(h_cancel)
summary(d_cancel)
summary(m_cancel)
```
For cancellations, I chose to run a regression for hour (h_cancel), day (d-Cancel)
and month (m_cancel). Hours appears to have a positive but very small insignificant
relationship on cancellations. 

Day appears to have a negative and insignificant relationship on cancellation

Month also appears to have a negative and insiginficant relationship with cancellation

```{r, echo=FALSE}
delay_dest <- fpa %>%
  dplyr::group_by(dest) %>%
  dplyr::summarise(
    max_delay = max(actual_delay, na.rm=TRUE),
    avg_delay = mean(actual_delay, na.rm=TRUE),
    min_delay = min(actual_delay,na.rm=TRUE),
    count = n()
  )
```

##Indicating relationship between destination and delays
```{r, echo=FALSE}
delay_dest_obs <- delay_dest %>%
  dplyr::arrange(avg_delay, dest)
```

```{r, echo=FALSE}
head(delay_dest_obs)
```
This give the 6 destinations with the least amount of delays. 

```{r, echo=FALSE}
dest_cancel<-glm(flights$canceled~flights$dest, family=binomial(link="probit"))
```

#Indicating Relationship between destinations and cancellations

```{r, echo=FALSE}
summary(dest_cancel)
```
Here I've run a regression of flights cancelled on destinations. All destinations appear
to have a positive but very insignificant relationship with cancellations.
We can interpret that destinations will not really determine whether there is a cancellation. 

##Indicating relationship between plane characteristics and Delays

```{r, echo=FALSE, warning=FALSE}
delay.type<-ggplot(fpa,
                    aes(x=type, y=dep_delay))
delay.type + geom_point(aes(colour=origin, alpha=0.5))
```
From this we can interpret that the Fixed wing multi engine leads to the most delays compared to the fixed wing singe engine and the Rotorcraft. 


## Indicating relationship between plane characteristics and Cancellations
```{r, echo=FALSE, warning=FALSE}
cancel.type<-ggplot(fpa,
                   aes(x=seats, y=engine))
cancel.type + geom_point(aes(colour=cancel, alpha=0.5))
```
Blue dots are indicated by a flight being cancelled and black dots are indicated
by flights not being cancelled. From here we can see that a flight is more likely to be 
cancelled if it has fewer seats and is a turbo-jet plane or a turbo-fan plane. 